Probabilistic Multiclass Kernel Embedding Classifier

### 2017-04-16 kernel_embedding_classifier.tex
Add unpolished discussion about why SGD works

		We begin by selecting a mini-batch size $n_{b} << n$ that is much smaller than $n$. The objective \eqref{eq:trace_norm_optimisation} is then computed only for the mini-batch

		For gradient descent, pick $n_{b} = n$ so that $\mathcal{I} = \mathbb{N}_{n}$ and the full dataset is used to compute the gradients.

		SGD Coefficient ratio for gradient update can be absorbed into the learning rate so don't need to worry about it

		Also learning rate can be left to the same

		Convergence can usually be left to the same gradient error tolerance

		\textbf{Why does SGD work?} It is not immediately clear to me why SGD works. What we can see from experiments is that the batch data complexity bound is usually a biased estimator of the full data complexity bound. This bias is expected for one primary reason. The first is that this quantity is a data complexity bound to begin with, and thus while the actual batch data complexity may be an unbiased estimator of the full data complexity, the tightness of the upper bound is expected to change with growing datasets and thus this bias comes from the varying degrees of tightness in the bound as data increases. While this bias exists, we do see from experiments that, as we optimise the hyperparameters, both the batch data complexity bound and full data complexity bound decreases at a proportional pace. At first it may seem that there is simply a multiplicative factor (possibly in log scale or square root scale with respect to the data size) relationship between them, although it is unclear how this multiplicative factor can be computed. We did try seeing if the average complexity bound of disjoint batches is the complexity bound of the full batch. The answer is no but it is close, through numerical simulation. However, again this should be expected as the tightness of the bound is expected to change for varying data size. Perhaps the complexity itself is additive.

		Idea: Last time in my derivation I can derive that the full complexity bound is the sum of the batch complexity bound plus some co-varying terms (which can be negative). If I can show that the magnitude of such co-varying terms are upper bounded by the batch complexity bounds, then I can say that as the batch complexity bounds are minimised, so is the full complexity bound!

		We know that the SGD works because the gradients computed from each batch is consistently vanishing as we optimise the hyperparameters. Surprisingly, it seems to be on the same scale of the full GD case and suggests that the batch gradient is possibly a (scaled) consistent estimator of the full gradient as well.

		Furthermore, not only does SGD reduce the time and space complexity of each training iteration drastically, it does not increase the number of iterations required for convergence either (at least for a Gaussian kernel), and, in many cases actually takes less iterations than the full GD solution. This could be because the SGD gradients are usually under-biased compared to the full GD gradients, and thus reach the gradient error tolerance faster.

@@

		Round Shifts: Instead of taking completely random batches from the full dataset each time, what one can do instead is to use a ``round shift'' setup to choose the batches. That is, we first choose a batch size $n_{b}$. The first batch is the first $n_{b}$ data points in the dataset, the second batch is the second $n_{b}$ data points in the dataset, and so on. When the whole dataset is scanned through once, we repeat and start from the beginning again. Experiments show that this converges to basically the same solution with the full GD approach and also the vanilla SGD approach. However, it does take longer iterations to reach the solution. Our guess is that the data points we use each time in our batch complexity bound estimation is not independent from each other enough such that the gradients are more correlated to each other and is thus less adaptive.

>>>>

		We begin by selecting a mini-batch size $n_{b} << n$ that is much smaller than $n$. At each training iteration, we sample a mini-batch of size $n_{b}$ from the full dataset. The objective \eqref{eq:trace_norm_optimisation} and hence its gradients with respect to each hyperparameter is then computed only for the mini-batch. This stochastic gradient is then used to make hyperparameter updates in that training iteration. This continues until the magnitude of the stochastic gradients fall below a certain range. This training algorithm is presented in \cref{alg:training_stochastic_gradient_descent}. This implementation also generalises the gradient descent approach, where $n_{b} = n$ so that $\mathcal{I} = \mathbb{N}_{n}$ and the full dataset is used to compute the gradients.

		\note{I have written some not-so-polished comments about why SGD works below.}

		\textbf{Why does SGD work?} It is not immediately obvious that a stochastic gradient descent approach would work for a kernel-based classifier. Unlike the cross-entropy loss, the complexity bound objective \eqref{eq:trace_norm_optimisation} is not expressed as a sum of terms each depending on one observation from the dataset. Even if it is, being a kernel-based classifier, each term of such a objective would still require the computation of an entire gram matrix and its regularised inverse. As such, taking a mini-batch implies that we are only taking a block portion of the entire gram matrix and its regularised inverse.

		What we can see from experiments is that the batch data complexity bound is usually a biased estimator of the full data complexity bound. This bias is expected for one primary reason. The objective \eqref{eq:trace_norm_optimisation} is a data complexity bound to begin with, and thus while the actual batch data complexity may be an unbiased estimator of the full data complexity, the tightness of the upper bound is expected to change with growing datasets and thus this bias comes from the varying degrees of tightness in the bound as data increases. While this bias exists, we do see from experiments that, as we optimise the hyperparameters, both the batch data complexity bound and full data complexity bound decreases at a proportional pace. At first it may seem that there is simply a multiplicative factor (possibly in log scale or square root scale with respect to the data size) relationship between them, although it is unclear how this multiplicative factor can be computed. We did try seeing if the average complexity bound of disjoint batches is the complexity bound of the full batch. The answer is no but it is close, through numerical simulation. However, again this should be expected as the tightness of the bound is expected to change for varying data size. Perhaps the complexity itself is additive.

		However, it is also not that unintuitive if we realise that each mini-batch is an independently sampled subset of the full dataset. This means that the complexity of the model learned from each mini-batch should not be too different from one another, as they are all data from the same probability distribution. This gives intuition of how each stochastic gradient is on average correct if we want to learn a low complexity model \textit{for a dataset of the same size as the mini-batch}. However, this may not translate to a low complexity model for the full dataset. In fact, we do expect that the mini-batch complexity bound to be smaller than the full dataset complexity bound, since models learned on less data tend to be simpler than models learned on more data. Nevertheless, intuitively, it also makes sense that a low complexity model for a small dataset would mean that the model is also of low complexity for a larger dataset, as long as the small dataset can capture, albeit to smaller degrees, many of the variations of the larger dataset. We achieve this by randomly shuffle the mini-batch being used in each iteration. If it does not capture all such variations, the learned model may be too simple when applied on the full dataset.

		We know that, experimentally, SGD works, because the gradients computed from each batch is consistently vanishing as we optimise the hyperparameters. Surprisingly, it seems to be on the same scale of the full GD case and suggests that the batch gradient is possibly a (scaled) consistent estimator of the full gradient as well.

		Furthermore, not only does SGD reduce the time and space complexity of each training iteration drastically, it does not increase the number of iterations required for convergence either (at least for a few simple kernels like the Gaussian kernel), and, in many cases actually takes less iterations than the full GD solution. This could be because the SGD gradients are usually under-biased compared to the full GD gradients, and thus reach the gradient error tolerance faster.

@@

		Round Shifts: Instead of taking completely random batches from the full dataset each time, what one can do instead is to use a ``round shift'' setup to choose the batches. That is, the first batch is the first $n_{b}$ data points in the dataset, the second batch is the second $n_{b}$ data points in the dataset, and so on. When the whole dataset is scanned through once, we repeat and start from the beginning again. Experiments show that this converges to basically the same solution with the full GD approach and also the vanilla SGD approach. However, it does take longer iterations to reach the solution. Our guess is that the data points we use each time in our batch complexity bound estimation is not independent from each other enough such that the gradients are more correlated to each other and is thus less adaptive.

		Optimisation Setup: We mentioned that it is possible that the expectation of the gradients of the SGD is a scaled version of the full GD gradients. In that case, what is commonly done is to multiply the SGD gradients by the correct ratio. However, here we choose to leave that out, as such a ratio can simply be absorbed into the learning rate. Also, this can be advantageous as such a ratio is usually greater than $1$ (in order to scale up the estimate to the same scale as the full GD case), and if we do multiply this ratio to the stochastic gradients, the learning rate should usually be decreased, in order to dampen out the stochasticity from the stochastic gradients and avoid divergence. From experiments we see that by not multiplying this ratio, the learning rate can usually be left to the same as the full GD approach. For the same reason, when we switch from full GD to SGD this way, the same gradient error tolerance can also be used.

		Proof Idea: I suspect that the full complexity bound is the sum of the batch complexity bound plus some co-varying terms (which can be negative). If I can show that the magnitude of such co-varying terms are upper bounded by the batch complexity bounds, then I can say that as the batch complexity bounds are minimised (randomly), so is the full complexity bound!
